Evaluating the Performance and Energy Efficiency of the COSMO-ART model system

### 2014-04-25 paper_enahpc14/Sec3:Environment-setup.tex
Manuel's changes

\textsc{COSMO-ART}  is compiled  using the  Extrae  compiler wrappers,
which automatically  instrument the Fortran code of  the model.  Next,
\textsc{COSMO-ART}  is  run on  the  nodes,  thus dissipating  certain
amount  of power.   These  nodes are  connected  to power  measurement
devices that account for the dissipated power/consumed energy and send
the power  data to the  tracing server.  The client,  meanwhile, sends
start/stop  primitives  in  order  to  gather  captured  data  by  the
watt-meters onto the  tracing server, where an instance  of the \pmlib
server is running.

Once  the run  is finished,  a file  containing the  power  profile is
created using the power data  received from the tracing server and the
instrumentation post-processing generates the performance trace files.
All  these files  are  combined then  in  Paraver which  allows us  to
visualise   the   performance  trace   and   the   power  profile   of
\textsc{Cosmo-art} all together.

@@

The \textsc{Cosmo-art}  baseline is a  pure MPI based  Fortran~90 code
currently  running  on   distributed  multi-core  systems  only.   The
software stack on the platforms was controlled using the modules
framework which gives an easy  and flexible mechanism to access to all
of the  provided compilers, tools  and applications.  For  our initial
benchmarking,   we  opted   for   the  GNU   compiler  gcc-4.8.1   on
\textsc{Monch} and gcc-4.8.2 on  \textsc{Pilatus} and \textsc{Tintorrum}) using the -O3 compiler
flag  in favor  of the  Intel compilers 14.0.1,  delivering inferior
performance.  In  addition, we installed the following implementations of MPI: MVAPICH2 1.9 on \textsc{Monch} and \textsc{Pilatus} and  OpenMPI 1.6.5 on \textsc{Tintorrum}. On the other hand, we  used HDF5 1.8.12 and NetCDF  (4.3.1) libraries for  the management  of extremely  large and
complex data collections.  All computes nodes have an operating system
based on GNU/Linux featuring ``2.6.32-358.11.1.el6.x86\_64'' kernel in
\textsc{Monch}     and      ``3.0.101-0.15-default''     kernel     in
\textsc{Pilatus}.

>>>>

\textsc{COSMO-ART}  is compiled  using the  Extrae  compiler wrappers,
which automatically  instrument the Fortran code of  the model.  Next,
\textsc{COSMO-ART}  is  run on  the  nodes,  thus dissipating  certain
amount  of power.   These  nodes are connected to power distribution units (PDUs)
APC8653 that account for the dissipated power of at 1\,Hz, these data is afterwards sent
to the  tracing server. The client,  meanwhile, sends start/stop  primitives  in  order  to  gather  captured  data  by  the watt-meters onto the  tracing server, where an instance  of the \pmlib server is running.

Once  the run  is finished,  a file  containing the  power  profile is
created using the power data  received from the tracing server and the
instrumentation post-processing generates the performance trace files.
All  these files  are  combined then  in  Paraver which  allows us  to
visualise   the   performance  trace   and   the   power  profile   of
\textsc{COSMO-ART} all together.

@@

The \textsc{COSMO-ART}  baseline is a  pure MPI based  Fortran~90 code
currently  running  on   distributed  multi-core  systems  only.   The
software stack on the platforms was controlled using the modules
framework which gives an easy  and flexible mechanism to access to all
of the  provided compilers, tools  and applications.  For  our initial
benchmarking,   we  opted   for   the  GNU   compiler  gcc-4.8.1   on
\textsc{Monch} and gcc-4.8.2 on  \textsc{Pilatus} and \textsc{Tintorrum}) using the -O3 compiler
flag  in favor  of the  Intel compilers 14.0.1,  delivering inferior
performance.  In  addition, we installed the following implementations of MPI: MVAPICH2 1.9 on \textsc{Monch} and \textsc{Pilatus} and  OpenMPI 1.6.5 on \textsc{Tintorrum}. On the other hand, we  used HDF5 1.8.12 and NetCDF  (4.3.1) libraries for  the management  of extremely  large and
complex data collections.  All computes nodes have an operating system
based on GNU/Linux featuring ``2.6.32-358.11.1.el6.x86\_64'' kernel in
\textsc{Monch}     and      ``3.0.101-0.15-default''     kernel     in
\textsc{Pilatus}.
